Geometric variational problems

### 2020-11-12 NSF2020/Description/specificaim1_ver2.tex
finishing up 4.2.2

\begin{wrapfigure}[12]{l}{0.32\textwidth}
\centerline{\includegraphics[width=0.32\textwidth]{Figures/SaddleSplayDiagram.pdf}}
\caption{\label{fig:saddle_splay} \footnotesize Determination of
  saddle splay modulus.}
\end{wrapfigure}
%
In either case there is a dissipation of the total potential \eqref{eq:total_poten}, and this stabilizes the evolving bilayer shape.
Therefore, we reconstruct an evolving monolayer dividing surface $\Sigma$ and director field ${\bf n}$ by
interpolating the particle centers and orientations. Using \eqref{ansatz3}, we calculate a continuum energy $\mathbcal{W}$ from the interpolated shapes.

@@

We propose to conduct calculations similar to the example of figure~\ref{fig:flattening} for other elastic moduli such as the effective twist modulus $\KT$.
Molecular dynamics investigations find a twist modulus about 1 \kBT \cite{LeVeWa14}, and here the specific non-equilibrium shape consists of a single layer of amphiphilic particles on a hydrophobic substrate as illustrated by Figure~\ref{fig:deformations}D. Having a nonzero twist requires nonzero tilt, because the surface gradient of the lipid directoy equals the second fundamental form, and is therefore symmetric, whenever tilt is zero locally. The twist deformation is a fully three-dimensional deformation. \S subsec:specific_aim_3 addresses outstanding implementation issues like three-dimensional boundary integral equation solvers.

>>>>

\begin{wrapfigure}[12]{l}{0.32\textwidth}
\centerline{\includegraphics[width=0.32\textwidth]{Figures/SaddleSplayDiagram.pdf}}
\caption{\label{fig:saddle_splay} \footnotesize Determination of
  saddle splay modulus.}
\end{wrapfigure}
%
In either case there is a dissipation of the total potential \eqref{eq:total_poten}, and this stabilizes the evolving bilayer shape.
Therefore, we reconstruct an evolving monolayer dividing surface $\Sigma$ and director field ${\bf n}$ by
interpolating the particle centers and orientations. Using \eqref{ansatz3}, we calculate a continuum energy $\mathcal{W}$ from the interpolated shapes.

@@

We propose to conduct calculations similar to the example of figure~\ref{fig:flattening} for other elastic moduli such as the effective twist modulus $\KT$.
Molecular dynamics investigations find a twist modulus about 1 \kBT
\cite{LeVeWa14}, and here the specific non-equilibrium shape consists of
a single layer of amphiphilic particles on a hydrophobic substrate as
illustrated by Figure~\ref{fig:deformations}D. Having a nonzero twist
requires nonzero tilt, because the surface gradient of the lipid
directoy equals the second fundamental form, and is therefore symmetric,
whenever tilt is zero locally. The twist deformation is a fully
three-dimensional deformation. \S\ref{subsec:specific_aim_3} addresses outstanding implementation issues like three-dimensional boundary integral equation solvers.

### 2020-11-12 NSF2020/Description/specificaim2_ver2.tex
finishing up 4.2.2

\subsubsection{Novel reciprocal identities}
In the course of our study, we have found it useful to devise certain reciprocal identities that
allow us to calculate forces accurately without having to resort to overly complicated numerical integration schemes. 
At issue is that there is a break-down in numerical accuracy for the gradient of water activity near the particle
boundaries (see Figure \ref{fig:bierror}).
This is unfortunate the boundary is exactly where we need the gradients to be accurate in order for the formulas \eqref{forceandtorque}
to be meaningful. To get around this issue, we prove the following identity
  \begin{equation}
    \label{eq:reciprocal}
{\bf F}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i}[\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S,\quad
{\bf G}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i} ({\bf x}-\mathbf{a}_i) \times [\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S
  \end{equation}
for $i = 1\dots, N$. Here $u = \sum_{i=1}^N u_i$ is a sum of solutions $u_i$ the screened Laplace equation,
$\boldsymbol{\sigma}_{ij} = \rho^{-1} u_iu_j {\bf I} + \rho(\nabla u_i \cdot \nabla u_j {\bf I} - 2 \nabla u_i \otimes \nabla u_j)$ and
$[\cdot]$ denotes the jump across $\partial \Omega.$  It turns out that because of \cite{eq:reciprocal},
we no longer require knowledge of the gradient on the boundary, which is enormously beneficial for calculating force and torque.

>>>>

\subsubsection{Novel reciprocal identities}
The force and torque formulas~\eqref{forceandtorque} require the
hydrophobic stress along the particle boundaries. Unfortunately,
standard quadrature formulas to estimate $\nabla u$ on the particle
boundaries introduces large amounts of quadrature error
(Figure~\ref{fig:bierror}). Therefore, it is useful to devise reciprocal
identities for the force and torque on body $i$ that does not
require integration along body $i$. This is achieved using the identity 
\begin{align}
    \label{eq:reciprocal}
{\bf F}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i}[\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S,\quad
{\bf G}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i} ({\bf
  x}-\mathbf{a}_i) \times [\boldsymbol{\sigma}_{ij} +
  \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S, 
\end{align}
for $i=1,\ldots,N$, which we have proved. Here $u = \sum_{i=1}^N u_i$ is
a sum of solutions $u_i$ the screened Laplace equation,
$\boldsymbol{\sigma}_{ij} = \rho^{-1} u_iu_j {\bf I} + \rho(\nabla u_i
\cdot \nabla u_j {\bf I} - 2 \nabla u_i \otimes \nabla u_j)$ and
$[\cdot]$ denotes the jump across $\partial \Omega.$  It turns out that
because of \eqref{eq:reciprocal}, we no longer require knowledge of the
gradient on the boundary, which is enormously beneficial for calculating
force and torque.
